Autocorrelation in Sparse Distributed Memory

### 2019-07-02 Chapters/sdm-autocorrelation.tex
macbookpro compiling
footnotes

\citet{Kanerva1994}

@@

@@

Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905$.  Each bit of a hard location does indeed have a small pull bias.  What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0---an intuition developed in Figure \ref{fig:bitmatchbias}.

\begin{figure}[!h]
\centering

@@

\begin{lemma}
Let $r$ be the access radius given that $f$ percent of the hard locations are activated. Then, $\lim_{n \rightarrow \infty} r/n = 1/2$.
\label{thm:sdm-access-radius}
\end{lemma}
\begin{proof}

As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 0.5, $B(n, 0.5)$. For $n$ sufficiently large, the Binomial distribution can be approximated by a Normal distribution, i.e., $B(n, 0.5) \rightarrow \mathcal{N}(\mu = n/2, \sigma^2 = n/4)$.

Let $\Phi(x)$ be the cdf of the standard normal distribution. Let $z = \frac{r - n/2}{\sqrt{n}/2}$. Thus, $P(d(x, y) \le r) = \Phi(z)$. As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$.

@@

@@

Another way to prove is to divide into two cases:

Suppose that $n$ is an even integer, then,

@@

In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has reduced significantly as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}. In fact, in this case, using Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = 0.516876$.

@@

So far we have looked only at a single pair of bitstrings, the probability of a single bitmatch between bitstrings within the access radius distance.  Now let us consider the number of activated hard locations exhibiting this bitmatch.

@@

Let $Z$ be the number of activated hard locations with the same bit as the reading address. Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.

@@

Applying the law of total variance, $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = p(1-p)\mu_h + p^2 \sigma^2_h$.
\end{proof}
As, in our case, $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.

See Figure \ref{fig:sdm-same-bit-histogram} for a comparison between the theoretical model and a simulation.

>>>>

Explain equator distance\footnote{explain equator distance}.

@@

\subsection{Bitmatches at $i$.}

@@

Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905$.  Each bit of a hard location does indeed have a small pull bias.  What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0 --- an intuition developed in Figure \ref{fig:bitmatchbias}.

TRANSITION\footnote{Transition to what's coming next.}

\begin{figure}[!ht]
\centering

@@

\begin{lemma}
Let $r$ be the access radius given that $f$ percent of the hard locations are activated. Then, $\lim_{n \rightarrow \infty} r/n = 1/2$\footnote{AS WRITTEN, this cannot be right.  This is saying that, as the number of dimensions explodes, o raio sobre o número de dimensões converge p/ 1/2?  Não, converge para zero.  Temos que re-escrever.}.
\label{thm:sdm-access-radius}
\end{lemma}
\begin{proof}

As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 1/2, $B(n, 1/2)$. For $n$ sufficiently large, the Binomial distribution can be approximated by a Normal distribution, i.e., $B(n, 1/2) \rightarrow \mathcal{N}(\mu = n/2, \sigma^2 = n/4)$.

Let $\Phi(x)$ be the cdf of the standard normal distribution. Let $z = \frac{r - n/2}{\sqrt{n}/2}$. \footnote{$z$ is, hence, measuring the number of standard deviations from the average in the Gaussian, as $r$ varies.}  Thus, $P(d(x, y) \le r) = \Phi(z)$. As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$\footnote{WTF?  $f$ isn't a percentage here, is it?}.

@@

so what?\footnote{WHAT DOES THIS MEAN?}

what now?\footnote{AND WHAT HAPPENS NOW?}

@@

\bigskip
Another way to prove is to divide into two cases:  Suppose that $n$ is an even integer, then,

@@

In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has reduced significantly---as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}. In fact, in this case, using Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = 0.516876$.

@@

So far we have looked only at a single pair of bitstrings: the probability of a single bitmatch between bitstrings within the access radius distance.  Now let us consider the number of activated hard locations exhibiting this bitmatch.

@@

Let $Z$ be the number of activated hard locations with the same bit as the reading address. Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.  \footnote{TRANSITION!!!}

@@

Applying the law of total variance, $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = p(1-p)\mu_h + p^2 \sigma^2_h$.
\end{proof}
As, in our case, $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.  Figure \ref{fig:sdm-same-bit-histogram} provides a comparison between the theoretical model and a simulation.
